---
title: "Compare EM accelaration methods on GMM model"
author: "DongyueXie"
date: "2022-12-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

Compare daarem,squarem,qusi-newton,pem.

## Set up environment

Load some packages and function definitions used in the example below.

```{r load-pkgs, warning=FALSE, message=FALSE}
library(ggplot2)
library(cowplot)
library(daarem)
library(turboEM)
source("../code/misc.R")
source("../code/mixem.R")
```

## Load data set

Load the 100,000 x 100 conditional likelihood matrix computed from a
simulated data set.

```{r load-data}
load("../data/mixdata.RData")
n <- nrow(L)
m <- ncol(L)
cat(sprintf("Loaded %d x %d data matrix.\n",n,m))
```

Set the initial estimate of the mixture proportions.

```{r init-estimate}
x0 <- rep(1/m,m)
```
## Run basic EM updates

```{r fit-em}
out <- system.time(fit1 <- mixem(L,x0,numiter = 200))
f1  <- mixobjective(L,fit1$x)
cat(sprintf("Computation took %0.2f seconds.\n",out["elapsed"]))
cat(sprintf("Log-likelihood at EM estimate is %0.12f.\n",f1))
```

## Run accelerated EM

```{r fit-daarem,eval=FALSE}
out <- system.time(fit2 <- mixdaarem(L,x0,numiter = 200))
f2  <- mixobjective(L,fit2$x)
cat(sprintf("Computation took %0.2f seconds.\n",out["elapsed"]))
cat(sprintf("Objective value at DAAREM estimate is %0.12f.\n",f2))
```

The maxiter in daarem and turboem seems to have different meanings. THe former one is the number of evaluations of fixptfn, the latter is not.

```{r}
out <- system.time(daarem_res <- daarem(x0,fixptfn=mixdaarem.update,objfn=mixdaarem.objective,L=L,e=1e-15,
                    control = list(maxiter=200,order=10,tol=1e-8,convtype='objfn')))
cat(sprintf("DAAREM took %0.2f seconds.\n",out["elapsed"]))
cat(sprintf("DAAREM obj is %0.12f.\n",daarem_res$value.objfn))
res = turboem(x0,fixptfn=mixdaarem.update,objfn=mixdaarem.objective,L=L,e=1e-15,method=c("squarem","pem","qn"),
              control.run = list(maxiter=200,convtype='objfn',keep.objfval=TRUE,tol=1e-8))
res
```


```{r}
f <- mixobjective(L,x)
plot(log(f-fit1$value),type='l',ylim=c(-2,10),ylab='distance to solution, log')
lines(log(f-daarem_res$objfn.track),col=2)
lines(log(f-res$trace.objfval[[1]]$trace),col=3)
lines(log(f-c(mixobjective(L,x0),res$trace.objfval[[2]]$trace)),col=4)
lines(log(f-c(mixobjective(L,x0),res$trace.objfval[[3]]$trace)),col=5)
legend('topright',c('em','daarem','squarem','pem','qn'),lty=c(1,1,1,1,1),col=1:5)
```
