---
title: "Compare EM accelaration methods on linear model, estimate prior"
date: 
site: workflowr::wflow_site
output: workflowr::wflow_html
---

 
```{r knitr, echo=FALSE}
knitr::opts_chunk$set(comment = "#",results = "hold",collapse = TRUE,
                      fig.align = "center")
```

Compared to [previous](fastEM_regression.html), here we also estimate the prior weights w.

## Analysis settings

These variables specify how the data are generated: `n` is the number
of simulated samples, `p` is the number of simulated predictors, `na`
is the number of simulated predictors that have a nonzero effect, `se`
is the variance of the residual.

```{r sim-settings}
n  <- 200 
p  <- 500
na <- 10
se <- 4
```

This specifies the prior on the regression coefficients: it is normal
with zero mean and variance `s0`.

```{r ridge-settings}
s0 <- 1/se
```

## Set up environment

Load some packages and function definitions used in the example below.

```{r load-pkgs, warning=FALSE, message=FALSE}
library(MASS)
library(daarem)
library(ggplot2)
library(cowplot)
library(turboEM)
theme_set(theme_cowplot())
source("../code/misc.R")
source("../code/datasim.R")
source("../code/mr_ash_prior.R")
```

Initialize the sequence of pseudorandom numbers.

```{r set-seed}
set.seed(1)
```

## Simulate a data set

Simulate predictors with "decaying" correlations.

```{r sim-x}
X <- simulate_predictors_decaying_corr(n,p,s = 0.5)
X <- scale(X,center = TRUE,scale = FALSE)
```

Generate additive effects for the markers so that exactly `na` of them have
a nonzero effect on the trait.

```{r sim-beta}
i    <- sample(p,na)
b    <- rep(0,p)
b[i] <- rnorm(na)
```

Simulate the continuous outcomes, and center them.

```{r sim-y}
y <- drop(X %*% b + sqrt(se)*rnorm(n))
y <- y - mean(y)
```


## Linear regression with mixture-of-normals priors

Next, we consider a less simple case in which the regression
coefficients are independently and identically drawn from mixture of
zero-centered normals; this can be seen as a multivariate extension to
the [adaptive shrinkage model][adaptive-shrinkage-paper], so we call
this "multivariate regression adaptive shrinkage" (mr-ash). Although
posterior computations with this model are more difficult than with
ridge regression, we can nonetheless obtain simple co-ordinate ascent
updates for computing posterior expectations of the coefficients if we
introduce a *variational approximation* to the posterior
distribution. The full derivation is omitted here; see the code in the
`mr_ash_update` function for details. (Note that the co-ordinate
ascent updates, unlike the ridge regression updates, are only
guaranteed to recover a *local maximum* of the objective function
being optimized.)

These two variables specify the variances and mixture weights for the
mixture-of-normals priors. Here we illustrate mr-ash with a prior
that is a mixture of three normals.

```{r mr-ash-settings}
b0 <- rep(0,p)
s0 <- c(0.1,1,10)^2/se
w  <- c(0.5,0.25,0.25)
k =length(w)
```

## Run mr-ash co-ordinate ascent updates

Fit the mr-ash model by running 100 iterations of the basic
co-ordinate ascent updates. 

```{r fit-mr-ash}
out <- system.time(fit3 <- mr_ash(X,y,b0,se,s0,w,numiter = 100))
cat(sprintf("Computation took %0.2f seconds.\n",out["elapsed"]))
fit3$value
```

## Run accelerated mr-ash co-ordinate ascent updates

```{r}
out <- system.time(daarem_res <- daarem(c(b0,w),fixptfn=fastEM_mr_ash_update,objfn=fastEM_mr_ash_objective,
                                        X=X,y=y,s=se,s0=s0,
                                        control = list(maxiter = 100,order = 10,tol = 0)))
cat(sprintf("daare took %0.2f seconds.\n",out["elapsed"]))
daarem_res$objfn.track

res = turboem(c(b0,w),fixptfn=fastEM_mr_ash_update,objfn=fastEM_mr_ash_objective,
                                        X=X,y=y,s=se,s0=s0,method=c("squarem","pem","qn"),
              control.run = list(maxiter=100,convtype='objfn',keep.objfval=TRUE,tol=0))
res
```

```{r}
f    <- max(c(fit3$value,daarem_res$objfn.track)) + 1e-8
plot(log(f-fit3$value),type='l',ylim=c(-20,6),ylab='distance to solution, log')
lines(log(f-daarem_res$objfn.track),col=2)
lines(log(f-res$trace.objfval[[1]]$trace),col=3)
lines(log(f-c(fit3$value[1],res$trace.objfval[[2]]$trace)),col=4)
lines(log(f-c(fit3$value[1],res$trace.objfval[[3]]$trace)),col=5)
legend('topright',c('em','daarem','squarem','pem','qn'),lty=c(1,1,1,1,1),col=1:5)
```


[daarem-paper]: https://doi.org/10.1080/10618600.2019.1594835
[adaptive-shrinkage-paper]: https://doi.org/10.1093/biostatistics/kxw041
